Thermodynamic fragments in magic state theories

### 2020-07-21 magic_paper.tex
bounds on Phi

Denote by $\Phi(m; n, p)$ the cumulative binomial distribution,
\begin{align}
	\Phi(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \\
	\Phi^c(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j}.
\end{align}
Note that $\Phi(m; n, p)^c$ differs only by a factor of $(-1)^j$.
It is clear that for any $n,p$ and $m \leq n$
\begin{align}
	&\Phi(m; n, p) + \Phi(n-m; n, 1-p) = 1, \label{eq:reversef}\\
	&\Phi^c(m; n, p) \leq \Phi(m; n, p),
\end{align}
where equality in the second line holds iff $m=0$.

@@

We now want to bound these points for any index $m$, so as to investigate the asymptotic behaviour of the Lorenz curve.
We define the upper bound
\begin{equation}
	U(m;n,p) \coloneqq 
	\begin{cases}
	e^{-nS\left(\frac{m}{n}||p \right)} \label{eq:ubd},&\ m \leq np, \\
	1,&\ m>np.
	\end{cases}
\end{equation}
\nick{we need to replace 1 with an exponential decay}
where 
\begin{equation}
	S(\alpha||\beta) \coloneqq \alpha \log{\frac{\alpha}{\beta}} + (1-\alpha) \log{\frac{1-\alpha}{1-\beta}},
\end{equation}
is the classical relative entropy.

There are two useful lower bounds for $\Phi(m;n,p)$, which we define as
\begin{equation}
	L(m;n,p) \coloneqq \max{\left\{ c(m;n) U(m;n,p), \\ 1 - U(n-m;n,1-p) \right\}} \label{eq:lbd},
\end{equation}
where
\begin{equation}
	c(m;n) = \left[8m\left(1-\frac{m}{n}\right)\right]^{-1/2}.
\end{equation}
Generally, the first expression within the $\max$ function is useful iff $m \leq np$, while the second expression is due to~\cref{eq:reversef} and thususeful iff $m \geq np$.
So we choose $L(m;n,p$ to be the maximum of the two at any $m$.
Figure~\cref{fig:Phi} demonstrates the upper and lower bounds for $\Phi(m;n,\frac{4}{5}$.

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phi.pdf}
    \caption{Distributions $\Phi(m;n,p)$ (blue) and $\Phi^c(m;n,p)$ (orange) for $n=30$ and $p=\frac{4}{5}$.
    The upper and lower bounds on $\Phi$ are also presented. 
    The upper bound is given by~\cref{eq:ubd} and the lower bound by~\cref{eq:lbd}.
    }
    \label{fig:Phi}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phi_nexp.pdf}
    \caption{Distributions $\Phi_+(m;n,p)/ U(m;n,p)$ (blue) and $U(m;n,p) \Phi_-(m;n,p)$ (orange) for $n=30$ and $p=\frac{4}{5}$.
    We multiply the bound in~\cref{eq:ubd} to extract the exponential behaviour of the two distributions.
    Here, $np = 24$, so for $m\geq 24$, $U(m;n,p) = 1$.
    }
    \label{fig:phinexp}
\end{figure}

>>>>

Denote by $\Phi(m; n, p)$ the cumulative binomial distribution,
\begin{align}
	\Phi(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phi}\\
	\Phi^c(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j}. \label{eq:phic}
\end{align}
Note that $\Phi(m; n, p)^c$ differs only by a factor of $(-1)^j$.
It is clear that for any $n,p$ and $m \leq n$
\begin{align}
	&\Phi(m; n, p) + \Phi(n-m; n, 1-p) = 1, \label{eq:reversef}\\
	&\Phi^c(m; n, p) \leq \Phi(m; n, p),
\end{align}
where equality in the second line holds iff $m=0$.

@@

We now want to bound these points for any index $m$, so as to investigate the asymptotic behaviour of the Lorenz curves.
We first define a useful function
\begin{equation}
	c(m;n) = 
	\begin{cases}
		0,&\ m=0, \\
		\left[8m\left(1-\frac{m}{n}\right)\right]^{-1/2},&\ 0<m<n, \\
		1,&\ m=n,
	\end{cases}
\end{equation}
as well as the classical relative entropy,
\begin{equation}
	S(\alpha||\beta) \coloneqq \alpha \log{\frac{\alpha}{\beta}} + (1-\alpha) \log{\frac{1-\alpha}{1-\beta}}.
\end{equation}

We now define two functions that we use to bound $\Phi(m;n,p)$,
\begin{align}
	U(m;n,p) &\coloneqq	e^{-nS\left(\frac{m}{n}||p \right)} \label{eq:ubd}, \\
	L(m;n,p) &\coloneqq c(m;n)e^{-nS\left(\frac{m}{n}||p \right)} \label{eq:lbd}.
\end{align}

\begin{lemma}\label{lem:ulbd}
	The cumulative distribution $\Phi(m;n,p)$ defined in~\cref{eq:phic} is exponentially bounded for all $m: 0\leq m \leq n$,
	\begin{equation}
		B_L(m;n,p) \leq \Phi(m;n,p) \leq B_U(m;n,p),\ m\leq np
	\end{equation}
	where
	\begin{equation}
		B_L(m;n,p) = \\
		\begin{cases}
			L(m;n,p),&\ m \leq np, \\
			1 - U(n-m;n,1-p),&\ m > np,
		\end{cases}
	\end{equation}
	and
	\begin{equation}
		B_U(m;n,p) =
		\begin{cases}
			U(m;n,p),&\ m \leq np, \\
			1 - L(n-m;n,1-p),&\ m > np.
		\end{cases}
	\end{equation}
\end{lemma}

The bounds are demonstrated in~\cref{fig:phi} for the interesting cases of $\Phi\left(m;n,\frac{1}{5}\right),\Phi\left(m;n,\frac{4}{5}\right)$.

\begin{figure}%
    \centering
    \subfigure[][]{%
    \label{fig:phi1}%
    \includegraphics[height=5cm]{figs/Phi.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\\%
    \subfigure[][]{%
    \label{fig:phi4}%
    \includegraphics[height=5cm]{figs/Phi2.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }
    \caption{Distributions $\Phi(m;n,p)$ (black) and $\Phi^c(m;n,p)$ (red) for $n=30$ and probability: 
    \subref{fig:phi1} $p=\frac{4}{5}$; \subref{fig:phi4} $p=\frac{1}{5}$.
    The upper and lower bounds on $\Phi(m;n,p)$ as stated in~\cref{lem:ulbd} are also presented. 
    }%
    \label{fig:phi}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phi_nexp.pdf}
    \caption{Distributions $\Phi_+(m;n,p)/U(m;n,p)$ (blue) and $\Phi_-(m;n,p)/U(m;n,p)$ (orange) for $n=30$ and $p=\frac{4}{5}$.
    We divide by the bound in~\cref{eq:ubd} to extract the exponential behaviour of the two distributions.
    Here, $np = 24$, so for $m\geq 24$, $U(m;n,p) = 1$.
    }
    \label{fig:phinexp}
\end{figure}
